Approximating the matrix exponential derivative of the CTMC generator matrix, with application to global SARS-CoV-2

### 2023-03-16 manuscript/main.tex

In addition to these scientific questions of interest, we are interested in the performance of the first-order approximation as a surrogate gradient for HMC in such a high-dimensional setting. To do isolate this performance, we temporarily fix the Bayesian bridge global-scale $\tau$ at 2.5$\times10^{-5}$.  We generate

>>>>

In addition to these scientific questions of interest, we are interested in the performance of the first-order approximation as a surrogate gradient for HMC in such a high-dimensional setting. To isolate this performance, we fix the Bayesian bridge global-scale $\tau$ at 2.5$\times10^{-5}$.  We generate a Markov chain with 80,000 states, saving every tenth state and removing the first 1000 states as burn-in.  Despite the relatively small number of iterations, we observe large ESS that suggest satisfactory accuracy of the first-order approximation within the context of high-dimensional surrogate HMC.  The ESS for the three fixed-effects regression coefficients $\theta_1$, $\theta_2$ and $\theta_3$ are 1,053, 1,343 and 498, respectively.  The median and minimum ESS for the 1,892 random-effects are 1,514 and 1,161, respectively.
